Simulations of Ballot Polling Risk-Limiting Audits

### 2022-10-12 providence/sections/workload.tex

Note there is also some constant overhead of workload for the whole audit, namely $C$ in Equation~\ref{eq:round_workload}, which we take to be zero in our examples but could be used by election officials to represent, for example, the effort of constructing a ballot manifest.
For simplicity, (and without loss of generality), we measure in multiples of the per ballot workload; that is, we assume it is one unit, $w_b=1$. A per round workload of $w_r=x$ corresponds to a per round workload which is $x$ times the per ballot workload. We use $w_r=1000$ as a conservative example. 
That is, we set the overhead of a round equal to the workload of sampling $1000$ ballots. Based on available data\cite{RI-report}, the time retrieving and analyzing each individual ballot is on the order of $75$ seconds which means that $w_r=1000$ is equivalent to roughly $20$ person-hours of workload. This corresponds to about $15$ minutes being spent, on average, per round in each of the $133$ counties of Virginia, a clearly conservative workload estimate. 
As shown in Figure~\ref{fig:with_round_workload}, lower average workloads are achieved by selecting higher stopping probability; \Providence achieves the lowest minimum average workload is achieved at roughly $0.7$.

\begin{figure}
\includegraphics[width=.5\textwidth]{with_round_workload.png}
\caption{For workload parameters $w_b=1$ and $w_r=1000$, this plot shows the expected workload for various round schedule parameters $p$. Expected workload is found using Equation~\ref{eq:workload} and the average number of ballots and rounds in our simulations as the expected number of ballots and rounds.}
\label{fig:with_round_workload}
\end{figure}

Importantly, this gives us a way to estimate the expected workload, as well as which round schedule value $p$ achieves it, for arbitrary round workload. For each round workload $w_r$, we produce a dataset analagous to that of Figure~\ref{fig:with_round_workload} and then find the minimum average workload achieved for each of the audits and its corresponding stopping probability $p$. 
Figure~\ref{fig:optimal_workloads} shows the optimal achievable workload for a wide range of per round workloads. For very low round workloads, the workload function approaches just the total number of ballots, and so, as seen in Figure~\ref{fig:abv_bals}, the three approaches differ by less. On the other hand, for extremely larger values of round workload, the average number of ballots has little impact on the workload function, and so the three audits again have similar values. For more reasonable values of the round workload $w_r$, SO \BRAVO and EoR \BRAVO achieve minimum workload roughly $1.1$ and $1.3$ times greater than that of \Providence.
Figure~\ref{fig:optimal_ps} shows the corresponding round schedule parameters $p$ that achieve these minimal workloads. As expected, a overhead for each round means that larger round sizes are needed to achieve an optimal audit, and so for all three audit $p$ increases as a function of $w_r$. Notice that \Providence is generally above and to the left of SO \BRAVO, and SO \BRAVO is generally above and to the left of EoR \BRAVO. This relationship reflects the fact that for the same round workload, \Providence can get away with a larger stopping probability because it requires fewer ballots.

\begin{figure}
\includegraphics[width=.5\textwidth]{optimal_workloads.png}
\caption{For varying round workload $w_r$, the optimal average workload achievable by each audit, as a fraction of the \Providence values. (We show the value for varying $w_r$ since the value of $w_r$ may differ significantly from case to case.) Note that values lower than $10^3$ are probably unrealistic in US statewide contests; in Virginia, $w_r=10^3$ corresponds to only about $15$ minutes of persontime per county as the overhead per round.}
\label{fig:optimal_workloads}
\end{figure}

\begin{figure}
\includegraphics[width=.5\textwidth]{optimal_ps.png}
\caption{The optimal (workload-minimizing) stopping probability $p$ for varying workload model parameters $w_r$. With $w_b=1$, the varying value of $w_r$ can equivalently be thought of as the ratio of the workload of a round to the workload of a ballot. (Note that the steps in this function are a consequence of our subsampling the workload function. That is, the workload-minimizing value of $p$ for each $w_r$ is only allowed to take on values at increments of $0.5$.)}
\label{fig:optimal_ps}
\end{figure}

\noindent\textbf{Precinct overhead.} For a more complete model, we can also introduce container-level workload. The time to sample a ballot from an entirely new box is typically greater than to sample a ballot from an already-open box. Based on a Rhode Island pilot RLA report\cite{RI-report}, this may mean that a ballot from a new container requires roughly twice the time as a ballot from an already-opened container. Typically available election results give per-precinct granularity of vote tallies, rather than individual container information. In Virginia, however, most precincts have a single ballot scanner whose one box has sufficient capacity for all the ballots cast in that precinct anyways, and so we model the per-container workload with an additional per-precinct constant workload, $w_p$. In this model, the workload estimate incurs an additional workload of $w_p$ every time a precinct is sampled from for the first time in a round. That is, let $E_{pi}$ be the expected number of distinct precincts sampled from in round $i$, and let $E_p=\sum_i E_{pi}$. Then the new model is
\begin{equation}
W(E_b, E_r, E_p) = E_b w_b + E_r w_r + E_p w_p + C
\label{eq:round_and_precinct_workload}
\end{equation}

@@

\subsection{Real time}
Given tight certification deadlines\footnote{Virginia recently passed legislation requiring pre-certification RLAs TODO check exactly.}, the total real time to conduct the RLA is also an important factor to consider when planning audits.
Because each county can sample ballots for the same round concurrently, the total real time for a round depends only on the slowest county. 
In Virginia, Fairfax County typically has the most votes cast by a significant difference; in the contest we consider, Fairfax County had ~551 thousand votes cast, more than double the ~203 thousand of second-highest Virginia Beach City.
Consequently, we model the expected total real time $T$ of an audit using just the largest county, and we define analagous variables for the expected values in just the largest county.
For the largest county, let the expected total ballots sampled be $\bar E_b$, the expected number of rounds $\bar E_r$, and the expected number of distinct precinct samples summed over all rounds be $\bar E_c$.
Similarly, we use real time per-ballot, per-round, and per-precinct workload variables, $t_b$, $t_r$, and $t_p$. So the real time of the audit is estimated by
\begin{equation}
T(\bar E_b, \bar E_r, \bar E_p ) = \bar E_b t_b + \bar E_r t_r + \bar E_p t_p + C
\label{eq:real_time}
\end{equation}

As before, we can use our simulations to estimate $\bar E_b$, $\bar E_r$, and $\bar E_p$ using the corresponding averages over the trials. 
Available data to estimate values for $t_b$, $t_r$, and $t_p$ is limited, and so we take as an example the values $t_b=75$ seconds, $t_r=3$ hours, and $t_p=75$ seconds.\footnote{The value $t_b=75$ seconds corresponds to a serial retrieval and interpretation of the ballots based on the \cite{RI-report} timing, $t_p=75$ seconds corresponds to the approximate doubling in time for new-box ballots as reported in \cite{RI-report} in the ballot-level comparison timing data, and $t_r=3$ hours is just a guess at an approximate order for this variable.} In practice, election officials could use our software and their own estimates of these values to explore choices for round schedules. Figure~\ref{fig:real_time} shows how the esimated real time for these values differs as a function of $p$. It should be noted that real values of $t_b$, $t_r$, and $t_p$ will vary greatly based on the number of parallel teams retreiving and checking ballots, the distribution of ballots and containers both in number and physical space, and other factors. We provide Figure~\ref{fig:real_time} only as an example of the general shape and behavior of this function. Use of this optimal scheduling tool would depend on parameter estimates tailored to each case.

>>>>

Note there is also some constant overhead of workload for the whole audit, namely $C$ in Equation~\ref{eq:round_workload}, which we take to be zero in our examples but could be used by election officials to represent, for example, the effort of constructing a ballot manifest.
For simplicity, (and without loss of generality), we measure in multiples of the per ballot workload; that is, we assume it is one unit, $w_b=1$. A per round workload of $w_r=x$ corresponds to a per round workload which is $x$ times the per ballot workload. We use $w_r=1000$ as a conservative example. 
That is, we set the overhead of a round equal to the workload of sampling $1000$ ballots. Based on available data\cite{RI-report}, the time retrieving and analyzing each individual ballot is on the order of $75$ seconds which means that $w_r=1000$ is equivalent to roughly $20$ person-hours of workload. This corresponds to about $15$ minutes being spent, on average, per round in each of the $133$ counties of Virginia, a clearly conservative workload estimate.

As shown in Figure~\ref{fig:with_round_workload}, average workloads first reduce as stopping probability increases; this is likely due to a decrease in the number of rounds. After hitting a sweet spot, average workloads again increase with stopping probability; this time, likely because the average number of rounds does not increase much and the cost changes because of number of ballots drawn, which increases with round size. \Providence achieves the lowest minimum average workload at roughly $p=0.7$ for our example choice of $w_r=1000$.

\begin{figure}[h!]
\includegraphics[width=.5\textwidth]{with_round_workload.png}
\caption{For workload parameters $w_b=1$ and $w_r=1000$, this plot shows the expected workload for various round schedule parameters $p$. Expected workload is found using Equation~\ref{eq:round_workload} and the average number of ballots and rounds in our simulations as the expected number of ballots and rounds.}
\label{fig:with_round_workload}
\end{figure}

Importantly, this gives us a way to estimate the minimum expected workload, as well as which round schedule value $p$ achieves it, for arbitrary round workload. For each round workload $w_r$, we produce a dataset analagous to that of Figure~\ref{fig:with_round_workload} and then find the minimum average workload achieved for each of the audits and its corresponding stopping probability $p$.

Figure~\ref{fig:optimal_workloads} shows the optimal achievable workload for a wide range of per round workloads. For very low round workloads, the workload function approaches just the total number of ballots, and so workload is minimized by minimizing the number of ballots drawn, which corresponds to small round sizes, and we would expect all three audits to behave similarly, as ballot-by-ballot audits, with the smallest workload. On the other hand, for extremely large values of round workload, the average number of ballots has little impact on the workload function, and so the three audits again have similar values, all corresponding to large round sizes in order to minimize the number of rounds.  We know that there is variation in the number of ballots used by each type of audit for large round sizes (a factor of two for $p=0.9$), but these values would be small in comparison to $w_r$. We observe this behaviour in Figure~\ref{fig:optimal_workloads} for extremely small and large workload values. For more reasonable values of the round workload $w_r$, SO \BRAVO and EoR \BRAVO achieve minimum workload roughly $1.1$ and $1.3$ times greater than that of \Providence.
\begin{figure}
\includegraphics[width=.5\textwidth]{optimal_workloads.png}
\caption{For varying round workload $w_r$, the optimal average workload achievable by each audit, as a fraction of the \Providence values.}
\label{fig:optimal_workloads}
\end{figure}

Figure~\ref{fig:optimal_ps} shows the corresponding round schedule parameters $p$ that achieve these minimal workloads. As expected, an overhead for each round means that larger round sizes are needed to achieve an optimal audit, and so for all three audits $p$ increases as a function of $w_r$. Notice that \Providence is generally above and to the left of SO \BRAVO, and SO \BRAVO is generally above and to the left of EoR \BRAVO. This relationship reflects the fact that for the same round workload, \Providence can get away with a larger stopping probability because it requires fewer ballots.

\begin{figure}
\includegraphics[width=.5\textwidth]{optimal_ps.png}
\caption{The optimal (workload-minimizing) stopping probability $p$ for varying workload model parameters $w_r$. (Note that the steps in this function are a consequence of our subsampling the workload function. That is, the workload-minimizing value of $p$ for each $w_r$ is only allowed to take on values at increments of $0.05$.)}
\label{fig:optimal_ps}
\end{figure}

\subsubsection{Precinct overhead.} For a more complete model, we can also introduce container-level workload. If. around requires multiple ballots from a single container, the container need only be unsealed once. Based on a Rhode Island pilot RLA report\cite{RI-report}, this may mean that a ballot from a new container requires roughly twice the time as a ballot from an already-opened container. Typically available election results give per-precinct granularity of vote tallies, rather than individual container information. In Virginia, however, most precincts have a single ballot scanner whose one box has sufficient capacity for all the ballots cast in that precinct anyways, and so we model the per-container workload as a per-precinct workload, $w_p$. In this model, the workload estimate incurs an additional workload of $w_p$ every time a precinct is sampled from for the first time in a round. That is, let $E_{pi}$ be the expected number of distinct precincts sampled from in round $i$, and let $E_p=\sum_i E_{pi}$. Then the new model is
\begin{equation}
W(E_b, E_r, E_p) = E_b w_b + E_r w_r + E_p w_p + C
\label{eq:round_and_precinct_workload}
\end{equation}

@@

\subsection{Real time}
Given tight certification deadlines\footnote{Virginia recently passed legislation requiring pre-certification RLAs.}, the total real time to conduct the RLA is also an important factor to consider when planning audits.
Because each county can sample ballots for the same round concurrently, the total real time for a round depends only on the slowest county. 
In Virginia, Fairfax County typically has the most votes cast by a significant difference; in the contest we consider, Fairfax County had ~551 thousand votes cast, more than double the ~203 thousand of second-highest Virginia Beach City.
Consequently, we model the expected total real time $T$ of an audit using just the largest county, and we define analagous variables for the expected values in just the largest county.
For the largest county, let the expected total ballots sampled be $\bar E_b$, the expected number of rounds $\bar E_r$, and the expected number of distinct precinct samples summed over all rounds be $\bar E_c$.
Similarly, we use real time per-ballot, per-round, and per-precinct workload variables, $t_b$, $t_r$, and $t_p$. So the real time of the audit is estimated by
\begin{equation}
T(\bar E_b, \bar E_r, \bar E_p ) = \bar E_b t_b + \bar E_r t_r + \bar E_p t_p + C
\label{eq:real_time}
\end{equation}

As before, we can use our simulations to estimate $\bar E_b$, $\bar E_r$, and $\bar E_p$ using the corresponding averages over the trials. 
Available data to estimate values for $t_b$, $t_r$, and $t_p$ is limited, and so we take as an example the values $t_b=75$ seconds, $t_r=3$ hours, and $t_p=75$ seconds.\footnote{The value $t_b=75$ seconds corresponds to a serial retrieval and interpretation of the ballots based on the \cite{RI-report} timing, $t_p=75$ seconds corresponds to the approximate doubling in time for new-box ballots as reported in \cite{RI-report} in the ballot-level comparison timing data, and $t_r=3$ hours is just a guess at an approximate order for this variable.} In practice, election officials could use our software and their own estimates of these values to explore choices for round schedules. Figure~\ref{fig:real_time} shows how the estimated real time for these values differs as a function of $p$. It should be noted that real values of $t_b$, $t_r$, and $t_p$ will vary greatly based on the number of parallel teams retrieving and checking ballots, the distribution of ballots and containers both in number and physical space, and other factors. We provide Figure~\ref{fig:real_time} only as an example of the general shape and behavior of this function. Use of this optimal scheduling tool would depend on parameter estimates tailored to each case.
